Supplementary Material for Large Linear Multi-output Gaussian Process Learning for Time Series

### 2017-05-14 paper/supplement.tex
Updated supplement to reflect truncated normal initialization

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{booktabs}
\usepackage{subfigure}

@@

\subsection{Representation evaluation}\label{empirical-rep}

@@

For each setting, we randomly sample entries for each $A_q,\bsk_q$, and $\bse$ and the inverse length scale $\gamma$ for each kernel. Then, we investigate the average gradient construction accuracy and speed of LLGP for different settings of $Q,\epsilon$, and \texttt{ktype}.Each kernel's inverse length scales and periods were selected by sampling uniformly in log space from 1 to 10 with $Q$ samples. Next, we construct a random LMC kernel by sampling entries of each $A_q$ from a standard normal distribution, $\bsk_q$ from an inverse gamma with unit shape and scale, and independent noise $\bse$ for every output from an inverse gamma with unit scale and mean $\epsilon$. Inputs and outputs were independently generated from $\Unif[0,1]$ for benchmarking.

@@

We next evaluated the accuracy of the gradients for $N=10$ trace samples. Fixing $R=3,n=5000,D=10$, we quantified the accuracy and speed of computing $\nabla\mcL$. Since, for each partial derivative, LLGP requires only $N$ $n$-sized vector dot products, it generally runs faster than the exact approach (Fig.~\ref{fig:rellog}), which must compute a matrix-matrix dot product. The gradients between the two, however, are virtually indistinguishable for smooth kernels that induce diagonally dominant covariance matrices (Fig.~\ref{fig:relgrad}). Kernels such as the single Mat\'{e}rn or periodic kernel with noise on the order of $10^{-4}$ lead to less accurate gradients, owing to poor \textsc{minres} convergence in the inversions (Fig.~\ref{fig:rellog}. We will show that the stochastic gradients suffice for optimization in practical examples.

\begin{figure}[!h]
  \centering

 \subfigure[time]{
    \resizebox{0.48\textwidth}{!}{\includegraphics{time_ratio.eps}}}
  \subfigure[inversion accuracy]{
    \resizebox{0.48\textwidth}{!}{\includegraphics{relalpha_l2.eps}}}
 \caption{Negative logarithm of the ratio of the exact algorithm to LLGP's analogue for (a) time and (b) accuracy metrics for evaluating gradient computation performance. In each case, higher is better, and extremal values are noted.  For each data point, the average was taken over five runs. Reducing average noise, corresponding to increases in $-\log_{10}\varepsilon$, and increases in $Q$, generally make gradient reconstruction more difficult by making $K$ more ill-conditioned, reducing the accuracy of \textsc{minres}.}
\label{fig:rellog}
\end{figure}

\begin{figure}[!h]
\centering
{\includegraphics[width=\textwidth]{relgrad_l2.eps}}
\caption{Negative logarithm of the relative error in $\nabla\mcL$ between LLGP and the exact computation, over the same grid of kernels as Fig.~\ref{fig:rellog}}\label{fig:relgrad}
\end{figure}

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{booktabs}
\usepackage[section]{placeins}

@@

\section{Representation evaluation}\label{empirical-rep}

@@

For each setting, we randomly sample entries for each $A_q,\bsk_q$, and $\bse$ and the inverse length scale $\gamma$ for each kernel. Then, we investigate the average gradient construction accuracy and speed of LLGP for different settings of $Q,\epsilon$, and \texttt{ktype}.Each kernel's inverse length scales and periods were selected by sampling uniformly in log space from 1 to 10 with $Q$ samples. Next, we construct a random LMC kernel by sampling entries of each $A_q$ from a standard normal distribution truncated to the unit interval, $\bsk_q$ from an inverse gamma with unit shape and scale, and independent noise $\bse$ for every output from an inverse gamma with unit scale and mean $\epsilon$. Inputs and outputs were independently generated from $\Unif[0,1]$ for benchmarking.

@@

We next evaluated the accuracy of the gradients for $N=10$ trace samples. Fixing $R=3,n=5000,D=10$, we quantified the accuracy LLGP's $\nabla\mcL$ by comparing against the exact Cholesky approach. The relative error in the gradients is generally low for smooth kernels that induce diagonally dominant covariance matrices (Fig.~\ref{fig:relgrad}). Kernels such as the single Mat\'{e}rn or periodic kernel with noise on the order of $10^{-4}$ lead to less accurate gradients, owing to poor \textsc{minres} convergence in the inversions (Fig.~\ref{fig:rellog}. We showed in the main paper that the stochastic gradients suffice for optimization in practical examples.

\begin{figure}[!h]
  \centering
{\includegraphics[width=\textwidth]{relgrad_l2.eps}}
\caption{Negative logarithm of the relative error from the LLGP gradient construction to the exact log likelihood gradient. In each case, higher is better, and extremal values are noted.  For each data point, the average was taken over five runs. Reducing average noise, corresponding to increases in $-\log_{10}\varepsilon$, and increases in $Q$, generally make gradient reconstruction more difficult by making $K$ more ill-conditioned, reducing the accuracy of \textsc{minres}.}
\label{fig:relgrad}
\end{figure}

\begin{figure}[!h]
\centering
{\includegraphics[width=\textwidth]{relalpha_l2.eps}}
\caption{Negative logarithm of the relative error in $K^{-1}\Ty$, using \textsc{minres} compared to the Cholesky solution. Higher is better, and extremal values are noted. An average was taken over five runs.}
\label{fig:rellog}
\end{figure}
